gewLesson 3
========================================================

***

### What to Do First?
Notes:
getwd()
list.files()
pf <- read.csv('pseudo_facebook.tsv', sep = '\t')
names(pf)
summary(pf)


***

### Pseudo-Facebook User Data
Notes:

```{r Pseudo-Facebook User Data}

```

***

### Histogram of Users' Birthdays
Notes:

```{r Histogram of Users\' Birthdays}
install.packages('ggplot2')
library(ggplot2)
qplot(x = dob_day, data = pf) +
  scale_x_continuous(breaks=1:31) +
  facet_wrap(~dob_month, ncol=3)

qplot(x = friend_count, data = pf, xlim = c(0, 1000))

qplot(x = friend_count, data = subset(pf, !is.na(gender)), binwidth = 10) +
  scale_x_continuous(limits =   c(0,1000),breaks = seq(0,1000,50))+
  facet_wrap(~gender, ncol=2)

qplot(x = friend_count, data = na.omit(pf)) + 
  facet_grid(gender ~ dob_month) 
  
table(pf$gender)  

by(pf$friend_count, pf$gender, summary)
```

***

#### What are some things that you notice about this histogram?
Response:

***

### Moira's Investigation
Notes:

***

### Estimating Your Audience Size
Notes:

***

#### Think about a time when you posted a specific message or shared a photo on Facebook. What was it?
Response:

#### How many of your friends do you think saw that post?
Response:

#### Think about what percent of your friends on Facebook see any posts or comments that you make in a month. What percent do you think that is?
Response:

***

### Perceived Audience Size
Notes:

***
### Faceting
Notes:

```{r Faceting}

```

#### Let’s take another look at our plot. What stands out to you here?
Response:

***

### Be Skeptical - Outliers and Anomalies
Notes:

***

### Moira's Outlier
Notes:
#### Which case do you think applies to Moira’s outlier?
Response:

***

### Friend Count
Notes:

#### What code would you enter to create a histogram of friend counts?

```{r Friend Count}

```

#### How is this plot similar to Moira's first plot?
Response:

***

### Limiting the Axes
Notes:

```{r Limiting the Axes}

```

### Exploring with Bin Width
Notes:

***

### Adjusting the Bin Width
Notes:

### Faceting Friend Count
```{r Faceting Friend Count}
# What code would you add to create a facet the histogram by gender?
# Add it to the code below.
qplot(x = friend_count, data = pf, binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50))
```

***

### Omitting NA Values
Notes:

```{r Omitting NA Values}

```

***

### Statistics 'by' Gender
Notes:

```{r Statistics \'by\' Gender}

```

#### Who on average has more friends: men or women?
mean vs median
long tail will pull the mean to left or right
median is a more robust statistic, as long as half of the data is trusted

Response:

#### What's the difference between the median friend count for women and men?
Response:

#### Why would the median be a better measure than the mean?
Response:

***

### Tenure
Notes:

```{r Tenure}
qplot(x = tenure/365, data = pf,
      xlab = 'Number of years using Facebook',
      ylab = 'Number of users in sample',
      color = I('black'), fill = I('#099DD9')) +
  scale_x_continuous(breaks = seq(1, 7, 1), limits = c(0, 7))


```

***

#### How would you create a histogram of tenure by year?

```{r Tenure Histogram by Year}

```

***

### Labeling Plots
Notes:

```{r Labeling Plots}

```

***

### User Ages
Notes:

```{r User Ages}
qplot(x=age, data = pf, binwidth = 1,
      xlab = 'Age of Facebook users',
      ylab = 'Number of Facebook users',
      color = I('black')) +
  scale_x_continuous(breaks = seq(0, 100,10), lim=c(0, 100))
```

#### What do you notice?
Response:

***

### The Spread of Memes
Notes:

***

### Lada's Money Bag Meme
Notes:

***

### Transforming Data
Notes:
```{r}
qplot(x = friend_count, data = pf)

summary(pf$friend_count)

summary(log10(pf$friend_count + 1))

summary(sqrt(pf$friend_count))
# define individual plots
p1 = ggplot(...)
p2 = ggplot(...)
p3 = ggplot(...)
p4 = ggplot(...)
# arrange plots in grid
grid.arrange(p1, p2, p3, p4, ncol=2)

# Add Log or Sqrt Scale to an Axis
scale_x_log10(...)
scale_y_log10(...)
scale_x_reverse(...)
scale_y_reverse(...)
scale_x_sqrt(...)
scale_y_sqrt(...)

```
```{r}
install.packages('gridExtra')
library(gridExtra)
p1 <- qplot(x = friend_count, data = pf)
p1
p2 <- qplot(x = log10(friend_count+ 1), data = pf) 
p2
p3 <- qplot(x = sqrt(friend_count), data = pf) 
p3
  
grid.arrange(p1, p2, p3, ncol=1)
p1 = ggplot(aes(x = friend_count), data = pf) + geom_histogram()
p2 = p1 + scale_x_log10()
p3 = p1 + scale_x_sqrt()

grid.arrange(p1, p2, p3, ncol=1)

```

***

### Add a Scaling Layer
Notes:

```{r Add a Scaling Layer}

```

***


### Frequency Polygons

```{r Frequency Polygons}
p1 = qplot(x = friend_count, y = ..count../sum(..count..),
      data = subset(pf, !is.na(gender)),
      xlab = 'Friend Count',
      ylab = 'Proportion of Users with that friend count',
      binwidth = 5, geom = 'freqpoly', color = gender) +
  scale_x_continuous(limits = c(200, 500), breaks = seq(0, 1000, 25)
      )




```

***

### Likes on the Web
Notes:

```{r Likes on the Web}
qplot(x = www_likes,
      data = subset(pf, !is.na(gender)),
      xlab = 'Web Count',
      ylab = 'Proportion of Users with that Web count',
      geom = 'freqpoly', color = gender) +
  
  scale_x_log10()
      

```


***

### Box Plots
Notes:

```{r Box Plots}
qplot(x = gender, y = log2(friend_count), 
      data = subset(pf, !is.na(gender) ), geom = 'boxplot')
```

#### Adjust the code to focus on users who have friend counts between 0 and 1000.

```{r}

```

***

### Box Plots, Quartiles, and Friendships
Notes:

```{r Box Plots, Quartiles, and Friendships}

```

#### On average, who initiated more friendships in our sample: men or women?
Response:
#### Write about some ways that you can verify your answer.
Response:
```{r Friend Requests by Gender}
qplot(x=gender, y = friendships_initiated, 
      data = subset(pf, !is.na(gender)), geom= 'boxplot') +
  coord_cartesian(ylim = c(0, 150))
```

Response:

***

### Getting Logical
Notes:

```{r Getting Logical}

summary(pf$mobile_likes)
summary(pf$mobile_likes > 0)
pf$mobile_check_in <- NA
pf$mobile_check_in <- ifelse(pf$mobile_likes > 0, 1, 0)
pf$mobile_check_in <- factor(pf$mobile_check_in)
summary(pf$mobile_check_in)
```

Response:
```{r}
sum(pf$mobile_check_in == 1)/length(pf$mobile_check_in)
```

***

### Analyzing One Variable
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!